\textbf{Dedukti: an implementation of set theory with pointed graphs} (titre provisoire)

### 2021-12-16 Article.tex
Add files via upload

\usepackage{amsmath}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz}
\usepackage{biblatex}
\usepackage{url}
\usepackage{fancyvrb}
\usepackage{enumitem}
\setitemize[1]{label=---}

@@

\begin{abstract}
We show how to implement set theory in \dedukti ~in such a way that it respects normalization. To do so, we adapt in $\lambda \Pi$\textit{--calculus modulo theory} the guideline given in \textit{Deduction modulo} by A. Miquel and G. Dowek --using a theory of pointed graphs. To achieve this goal, we develop a language of formulas.
\end{abstract}

@@

A previous work by A. Miquel and G. Dowek \cite{zermodulo} has demonstrated the implemtation of set theory in \textit{Deduction modulo} thaks to a structure of \textit{pointed graphs}. Since the writing of \cite{zermodulo}, \lpcm \ has been created as an improvement of \textit{Deduction modulo}. We adapt the implementation to the \lpcm \ and then implement it in \dedukti.

@@

To implement set theory in \dedukti, the first idea is to translate the axioms of the theory into rewriting rules. Nevertheless, as pointed out by M. Crabbé \cite{crabbé}, such implementation does not guarantee normalization. That is why A. Miquel and G. Dowek \cite{zermodulo} introduced a structure of pointed graphs to set theory. This set theory with pointed graphs -- called IZmod -- respects normalization.

@@

We saw in the examples that some different pointed graphs can represent the same set. To characterized this, A. Miquel and G. Dowek introduce bisimilarity between graphs (noted $\simeq$):

@@

Many definitions of the set theory intervines. The two classical set theories are Zermelo (Z) and Zermelo-Fraenkel (ZF). The set theory defined in \textit{Deduction modulo} by A. Miquel and G. Dowek \cite{zermodulo} is stronger than Zermelo set theory but weaker than Zermelo-Fraenkel set thery. Indeed, IZst contains Zermelo set theory, the Strong Extensionality axiom and the Transitive Closure axiom. \\

@@

The Strong Extensionality axiom implies the Extensionality axiom\footnote{See 5.3}.

@@

@@

We have implemented in \dedukti ~a version of set theory -- IZst -- that respects normalization properties. IZst set theory correponds to Zermelo set theory, with Strong Extensionality axiom and Transitive Closure axiom. To do so, we have adapted the work by A. Miquel and G. Dowek \cite{zermodulo} from \textit{Deduction modulo} to \lpcm. We have \textit{formally} written all the proofs of the lemmas allowing us to implement set theory in \dedukti.

>>>>

\usepackage{amsmath}
\usepackage[dvipsnames]{xcolor}
\usepackage{tikz}
\usepackage[backend=biber,style=alphabetic]{biblatex}
\usepackage{url}
\usepackage{fancyvrb}
\usepackage{enumitem}
\setitemize[1]{label=---}

@@

\begin{abstract}
We show how to implement set theory in \dedukti ~in such a way that it respects normalization. To do so, we adapt in $\lambda \Pi$\textit{--calculus modulo theory} the guideline given in \textit{Deduction modulo} by G. Dowek and A. Miquel --using a theory of pointed graphs. To achieve this goal, we develop a language of formulas.
\end{abstract}

@@

A previous work by G. Dowek and A. Miquel \cite{zermodulo} has demonstrated the implemtation of set theory in \textit{Deduction modulo} thaks to a structure of \textit{pointed graphs}. Since the writing of \cite{zermodulo}, \lpcm \ has been created as an improvement of \textit{Deduction modulo}. We adapt the implementation to the \lpcm \ and then implement it in \dedukti.

@@

To implement set theory in \dedukti, the first idea is to translate the axioms of the theory into rewriting rules. Nevertheless, as pointed out by M. Crabbé \cite{crabbé}, such implementation does not guarantee normalization. That is why G. Dowek and A. Miquel \cite{zermodulo} introduced a structure of pointed graphs to set theory. This set theory with pointed graphs -- called IZmod -- respects normalization.

@@

We saw in the examples that some different pointed graphs can represent the same set. To characterized this, G. Dowek and A. Miquel introduce bisimilarity between graphs (noted $\simeq$):

@@

Many definitions of the set theory intervines. The two classical set theories are Zermelo (Z) and Zermelo-Fraenkel (ZF). The set theory defined in \textit{Deduction modulo} by G. Dowek and A. Miquel \cite{zermodulo} is stronger than Zermelo set theory but weaker than Zermelo-Fraenkel set thery. Indeed, IZst contains Zermelo set theory, the Strong Extensionality axiom and the Transitive Closure axiom. \\

@@

The Strong Extensionality axiom implies the Extensionality axiom\footnote{See \ref{extensionality}}.

@@

\label{extensionality}

@@

We have implemented in \dedukti ~a version of set theory -- IZst -- that respects normalization properties. IZst set theory correponds to Zermelo set theory, with Strong Extensionality axiom and Transitive Closure axiom. To do so, we have adapted the work by G. Dowek and A. Miquel \cite{zermodulo} from \textit{Deduction modulo} to \lpcm. We have \textit{formally} written all the proofs of the lemmas allowing us to implement set theory in \dedukti.
